The Circumnuclear Media and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-02-09 ms.tex
salvaging the unsalvagable table2

Fig.~\ref{fig:cooling} shows the ratio of wind heating to radiative cooling, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:cooling2}] and surrounding discussion) as a function of radius for our fiducial solutions shown in Figure \ref{fig:profiles}.  For high heating rates of $v_{w} = 600$ and $1200$ km s$^{-1}$ cooling is unimportant across all radii, while for $v_{w} = 300$ km s$^{-1}$ we see that $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| < 1$ and $\tcool/\tff \lesssim 1,$ depending on the wind mass loss parameter $\eta$.  Note that the ratio $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ at the stagnation radius deviates from its minimum value across the entire grid by less than a factor of two; thus the ratio of $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ indeed represents a useful proxy for the presence of thermal instabilty across the entire grid.  Also note that for $v_{w}=600$ and $1200$ km s$^{-1}$ we have that $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| \sim t_{\rm cool}/t_{\rm ff}$ near the stagnation radius.  However, the two ratios diverge from each other at low heating ($v_{w} \lesssim 300$ km s$^{-1}$) because equations (\ref{eq:mdot_gas}), (\ref{eq:rhors}) underestimate the gas density, which is greater in a subsonic inflow (weak heating) instead of a supersonic outflow (strong heating).

@@

Key properties of the flow, such as the stagnation radius, the SMBH accretion rate, and the likelihood of thermal instability depend sensitively on the assumed heating rate $\propto qv_{w}^{2}$.  This section summarizes individual heating sources, providing estimates of the value of
\begin{align}
  v_{w} = \sqrt{\frac{2 t_h e}{\eta \rhostar}}
  \label{eq:vw_eff}
\end{align}
for each, where $e$ is volumetric heating rate.

@@

Energy and mass input to the CNM by stellar winds is the sum of contributions from main sequence and post-main sequence stars.  At early times following star formation, energy input is dominated by core collapse supernovae and by the fast outflows of Wolf-Rayet stars, while at later times energy input is dominated by main sequence winds.  Mass input is also dominated by massive stars for very young stellar populations, but for most stellar ages mass input is dominated by the slow AGB winds of low mass stars.

Appendix \ref{app:windheat} provides a calculation of wind heating rate from stellar winds, $v_{\rm w}^{\star}$, and the mass loss parameter, $\eta$ (eq.~[\ref{eq:q}]), as a function of age $\tau_{\star}$ of a stellar population which is formed impulsively (Figs.~\ref{fig:vwimp}, \ref{fig:etaImp}).  At very early times after the starburst ($\tau_{\star} \lesssim 10^{7}$), the wind heating rate exceeds 1000 km s$^{-1}$, while at much later times ($\tau_{\star} \sim t_{\rm h}$) stellar wind heating dominated by main sequence winds is much lower, $v_{\rm w} \sim 50-100 $ km s$^{-1}$ (see also \citealt{NaimanSoares-Furtado+:2013a}).  As we describe in $\S$, the net heating rate produced by stellar winds in the case of quasi-continuous star formation, as representative of the average star formation history of low mass galaxies, can also be significant, $v_{\rm w}^{\star} \gtrsim 1000$ km s$^{-1}$.  Thus stellar winds contribute a potentialy important source of energy as well as mass for the CNM.

@@

What is the effect of the Ia radius?  Between subsequent supernovae, stars release a gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$ interior to the Ia radius, which is gravitationally bound to the SMBH by an energy $E_{\rm bind} \sim M_{\rm g}\sigma^{2}$.  Thus, according to the above definitions, $E_{\rm Ia}/E_{\rm bind} \sim (v_{w}^{\rm Ia})^{2}/2\sigma^{2} \gtrsim 1$.  Hence Ia SN can in principle dynamically clear out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$.  Thus, even in cases when $r_{\rm s} > r_{\rm Ia}$, the SMBH accretion rate is limited to the value set by substituting the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}]) for $r_{\rm s}$ in the derivation of equation $\ref{eq:eddr_analytic}$, i.e.  
\begin{eqnarray}
\left.\frac{\dot{M}}{\dot{M}_{\rm edd}}\right|_{\rm max} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{R_{\rm Ia}}{R_{\rm soi}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    4.5 \times 10^{-4} M_{\bullet,8}^{-1.33}(\tau_{\star}/t_{\rm h})^{-0.2}
   & \text{core} \\
    2.2 \times 10^{-4} \Mbheight^{-0.84}(\tau_{\star}/t_{\rm h})^{-0.6}   & \text{cusp}.
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}
{\bf AG: Changed this expression somewhat from that of BDM (slightly
  different scaling relations were used?). I will
  check again! Note that the time dependence of eta~t$^{-1.3}$ is
  folded in above.}

>>>>

Fig.~\ref{fig:cooling} shows the ratio of wind heating to radiative cooling, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:cooling2}] and surrounding discussion) as a function of radius for our fiducial solutions shown in Figure \ref{fig:profiles}.  For high heating rates of $v_{w} = 600$ and $1200$ km s$^{-1}$ cooling is unimportant across all radii, while for $v_{w} = 300$ km s$^{-1}$ we see that $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| < 1$ and $\tcool/\tff \lesssim 1,$ depending on the wind mass loss parameter $\eta$.  Note that the ratio $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ at the stagnation radius deviates from its minimum value across the entire grid by less than a factor of two; thus the ratio of $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ indeed represents a useful proxy for the presence of thermal instabilty across the entire grid.  Also note that for $v_{w}=600$ and $1200$ km s$^{-1}$ we have that $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| \sim t_{\rm cool}/t_{\rm ff}$ near the stagnation radius.  However, the two ratios diverge from each other at low heating ($v_{w} \lesssim 300$ km s$^{-1}$) because equations (\ref{eq:mdot_gas}), (\ref{eq:rhors}) underestimate the true gas density, which is greater in a subsonic inflow (weak heating) than in a supersonic outflow (strong heating).

@@

Key properties of the flow, such as the SMBH accretion rate and the likelihood of thermal instability depend sensitively on the assumed heating rate $\propto qv_{w}^{2}$.  This section summarizes individual heating sources, providing estimates of the value of
\begin{align}
  v_{w} = \sqrt{\frac{2 t_h e}{\eta \rhostar}}
  \label{eq:vw_eff}
\end{align}
for each, where $e$ is volumetric heating rate.

@@

Energy and mass input to the CNM by stellar winds is the sum of contributions from main sequence and post-main sequence stars.  At early times following star formation, energy input is dominated by core collapse supernovae and by the fast outflows of Wolf-Rayet stars (e.g., \citealt{VossDiehl+:2009a}), while at later times energy input is dominated by main sequence winds (e.g., \citealt{NaimanSoares-Furtado+:2013a}).  Mass input is also dominated by massive stars for very young stellar populations, but for most stellar ages mass input is dominated by the slow AGB winds of evolved low mass stars.

Appendix \ref{app:windheat} provides a calculation of wind heating rate from stellar winds, $v_{\rm w}^{\star}$, and the mass loss parameter, $\eta$ (eq.~[\ref{eq:q}]), as a function of age $\tau_{\star}$ of a stellar population which is formed impulsively (Figs.~\ref{fig:vwimp}, \ref{fig:etaImp}).  At very early times after the starburst ($\tau_{\star} \lesssim 10^{7}$ yr), the wind heating rate exceeds 1000 km s$^{-1}$, while at much later times ($\tau_{\star} \sim t_{\rm h}$) stellar wind heating dominated by main sequence winds is much lower, $v_{\rm w} \sim 50-100 $ km s$^{-1}$.  As we describe in $\S$, the net heating rate produced by stellar winds in the case of quasi-continuous star formation, as representative of the average star formation history of low mass galaxies, can also be significant, $v_{\rm w}^{\star} \gtrsim 1000$ km s$^{-1}$.  Thus stellar winds contribute a potentialy important source of energy as well as mass for the CNM.

@@

Another effect of SN Ia is that they cap the accretion rate onto the SMBH by periodically blowing gas out of the nucleus of low mass galaxies.  Between subsequent supernovae, stars release a gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$ interior to the Ia radius, which is gravitationally bound to the SMBH by an energy $E_{\rm bind} \sim M_{\rm g}\sigma^{2}$.  Thus, according to the above definitions, $E_{\rm Ia}/E_{\rm bind} \sim (v_{w}^{\rm Ia})^{2}/2\sigma^{2} \gtrsim 1$.  Hence Ia SN can in principle dynamically clear out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$, especially for low mass BHs with $\sigma \ll v^{\rm Ia}_{w}$.  Thus, even in cases when $r_{\rm s} > r_{\rm Ia}$, the accretion rate onto low mass SMBHs is limited to the value set by substituting the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}]) for $r_{\rm s}$ in the derivation leading to our estimate of $\dot{M}$ (eq.~[$\ref{eq:eddr_analytic}$]), i.e.  
\begin{eqnarray}
\frac{\dot{M}_{\rm max,Ia}}{\dot{M}_{\rm edd}} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{R_{\rm Ia}}{R_{\rm soi}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    4.5 \times 10^{-4} M_{\bullet,8}^{-1.33}(\tau_{\star}/t_{\rm h})^{-0.2}
   & \text{core} \\
    2.2 \times 10^{-4} \Mbheight^{-0.84}(\tau_{\star}/t_{\rm h})^{-0.6}   & \text{cusp}.
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}
Note that in general $\dot{M}_{\rm max,Ia}$ exceeds the maximum accretion rate for a thermally stable flow $\dot{M}_{\rm max,TI}$ (eq.~[\ref{eq:Mdotmax}]), implying that dynamical `blow out' from SNe Ia cannot alone prevent cooling instabilities in low mass galaxies.

{\bf AG: Changed this expression somewhat from that of BDM (slightly
  different scaling relations were used?). I will
  check again! Note that the time dependence of eta~t$^{-1.3}$ is
  folded in above.}
